---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Qilin Zhang"
date: "2023/1/26"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

The data has 6 columns and 1643 rows. Each row represents the data for an individual dennys store. There are six variables which are address, city, state, zip, longitude, latitude.

```{r dim_dennys}
dim(dennys)
str(dennys)
```

### Exercise 2

The data has 6 columns and 909 rows. Each row represents the data for an individual laquinta location. There are six variables which are address, city, state, zip, longitude, latitude.

```{r dim_laquinta}
dim(laquinta)
str(laquinta)

```

### Exercise 3

There are a lot of laquinta location outside the US. There are 76 in Mexico, three in Turkey, two in China and canada, one in New Zealand, United Arab Emiratesm, Chile, and Colombia.

There are also a lot of dennys location outside the us. There are "one in Chile, 60 in Canada, five in Mexico, three in Costa Rica, two in Honduras, two in El Salvador, two in the United Kingdom and eight in New Zealand."
source: https://en.wikipedia.org/wiki/Denny%27s#:~:text=1%2C593%20of%20Denny's%201%2C698%20restaurants,and%20eight%20in%20New%20Zealand.

### Exercise 4

We can try to restrict the latitude and longitude data based on the range of US territory and filter out all the latitude and longitude combination that are not in the range. 
We can also look at the states data. If a location is not in the US, it will likely contain state data from the US. 

### Exercise 5

There are no location outside the US in this data set.

```{r dennys_outside}
dennys %>%
  filter(!(state %in% states$abbreviation)) 

```

### Exercise 6

```{r dennys_country}

dn <- dennys %>%
  filter((state %in% states$abbreviation)) %>%
  mutate(country = "United States")
  
```

### Exercise 7

```{r}
lq <- laquinta 

laquinta%>%
  filter(!(state %in% states$abbreviation)) %>%
  select(state) %>%
  print()

```

### Exercise 8

```{r Country}
lq <- laquinta %>%
  mutate(country = case_when(
    state %in% state.abb     ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT"           ~ "Colombia",
    state == c("AG","QR","CH","NL","VE","PU","SL")    ~ "Mexico",
    state == "FM"           ~ "Micronesia"
  )
)

lq %>%
  filter(!(country == "United States")) %>%
  select(country) %>%
  print()

lq <- lq %>%
  filter(country == "United States")
```

### Exercise 9

```{r state_count}
dn %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  arrange(desc(n))
```

### Exercise 10

DC has the most Denny's per thousand square mile.

RI has the most la Quinta per thousand square mile.
```{r}

dn %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  mutate(nmi2= area/n)%>%
  arrange(nmi2)

lq %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  mutate(nmi2= area/n)%>%
  arrange(nmi2)

dn <- dn %>%
  mutate(establishment = "Denny's")
lq <- lq %>%
  mutate(establishment = "La Quinta")

dn_lq <- bind_rows(dn, lq)

ggplot(dn_lq, mapping = aes(x = longitude,
                            y = latitude,
                            color = establishment)) +
  geom_point(alpha = 0.3)
```

### Exercise 11

I think his joke holds here

```{r NC_plot}

dn_lq %>%
  filter(state == "NC")%>%
  ggplot(dn_lq, mapping = aes(x = longitude,
                            y = latitude,
                            color = establishment)) +
  geom_point(alpha = 0.5)

```

### Exercise 12

I think his joke holds here. There seems like not much of a deviation between their locations.

```{r TX_plot}

dn_lq %>%
  filter(state == "TX")%>%
  ggplot(dn_lq, mapping = aes(x = longitude,
                            y = latitude,
                            color = establishment)) +
  geom_point(alpha = 0.2)


```

